Bird Audio Detection Challenge

### 2017-03-17 paper/main.tex
Added figures and finished most of writing

@@

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for CVPR Proceedings}

@@

%%%%%%%%% ABSTRACT
\begin{abstract} TODO: Make Abstract last \end{abstract}

@@

%-------------------------------------------------------------------------
\subsection{Data preparation and augmentation}

@@

We also deployed data augmentation techniques. Given one base sound clip,
positive or negative, and another negative sound clip, we concatenate them
along the y-axis, with the base clip weighted by a factor \(r\) between 0.25
and 1, and the negative sound clip weighted by a factor \(1-r\). So we
have:
\begin{equation}
	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
\end{equation}
If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
equals to 0.25, only 25\% of the base sound clip will be preserved. We
prevent \(r\) from going below 0.25 to ensure bird sound is still
detectable after augmentation.

@@

Network also has shortcuts from earlier layers of the network to later
stages, allowing easier gradient back-propagation to early layers of the
network. See Table 1 for network details.

@@

@@

\subsection{Method 3}

This method is the winning algorithm, submitted by Thomas Grill.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

>>>>

\graphicspath{ {figures/} }

\newenvironment{alt_enumerate}
{ \begin{enumerate}
		\setlength{\itemsep}{1pt}
		\setlength{\parskip}{1pt}
		\setlength{\parsep}{1pt} }
	{ \end{itemize} }

@@

%%%%%%%%% TITLE
\title{Bird Audio Detection Challenge}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}

	Acoustic detection of birds in the wild is useful in many biological
	scenarios, including estimation of bird population and species-specific
	acoustic monitoring. Most of the current methods are either manual or
	semiautomatic, requiring labor-intensive processing/tuning. We hereby
	present three alternative methods, as part of the Bird Audio Detection
	Challenge, hosted by Queen Mary University of London. Two methods are of
	our own submission, with the third one as the winning solution submitted
	by Thomas Grill. All methods utilize deep learning for automatic,
	tuning-free bird presence detection.

\end{abstract}

@@

%-------------------------------------------------------------------------
\subsection{Data preparation}

@@

@@

Network also has shortcuts from earlier layers of the network to later
stages, allowing easier gradient back-propagation to early layers of the
network. See Figure 1 for network details.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.5\textwidth]{v5_diagram}
	\caption{Network v5 Architecture. Content with box refers to kernel size,
	number of filters, and stride size}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{v8_diagram}
	\caption{Network v8 Architecture. Content of convolutional blocks refer
	to kernel size, number of filters, and stride size. Content of avgpool
	blocks refer to kernel size and stride size.}
\end{figure*}

%------------------------------------------------------------------------
\subsection{Method 3}

% TODO: Explain that spectrograms were used during training

This method is the winning algorithm of the challenge, submitted by Thomas
Grill. The neural network used consists of 4 convolutional layers with
pooling, and 2 dense layers. Multiple signal pre-processing techniques were
used, including silence/noise trimming, and a different data augmentation
technique involving shifting the spectra in time and frequency.
Pseudo-labeling and bagging techniques are also deployed, which we will
discuss in section 4.

%------------------------------------------------------------------------
\section{Experiments}

During our development and training of the network, we experimented with
various pre- and postprocessing techniques, as well as some training
strategies.

%------------------------------------------------------------------------
\subsection{Implementation Details}

The given labeled data from both dataset is split into 10 folds. Within
each fold, number of positive and negative samples are balanced. During
training, number of samples loaded from each dataset are also balanced, to
ensure no bias.

For in-house testing, 9 folds are used for training and 1 folf for testing.
For submission, all 10 folds are used for training.

For our method 1 and 2, we implemented them in python with Tensorflow
machine learning library. We trained from scratch, with learning rate of
0.03, learning rate decay by a factor of 0.1 every 40000 iterations, and a
total of 300000 iterations.

%------------------------------------------------------------------------
\subsection{Data Augmentation}

We deployed data augmentation as our only pre-processing technique. Given
one base sound clip, positive or negative, and another negative sound clip,
we concatenate them along the y-axis, with the base clip weighted by a
factor \(r\) between 0.25 and 1, and the negative sound clip weighted by a
factor \(1-r\). So we have:
\begin{equation}
	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
\end{equation}
If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
equals to 0.25, only 25\% of the base sound clip will be preserved. We
prevent \(r\) from going below 0.25 to ensure bird sound is still
detectable after augmentation.

%------------------------------------------------------------------------
\subsection{Reducing capacity}

One main difficulty we faced is the small size of training data. With no
regularization, our networks consistently over-fit to the training data,
perform near perfect on training and our own hold-out test data, but
perform poorly on the challenge data. To compensate for the lack of data,
we reduce the capacity of our models by reducing the number of filters for
each convolutional layer. We do so until the network cannot perform
perfectly on the training and testing data.

%------------------------------------------------------------------------
\subsection{Prediction Bagging}

This post-processing technique averages prediction scores from many models.
Due to the random cropping of audio signals during the data loading phase,
different evaluation of the same audio clip will yield slightly different
results. Therefore we predict 100 sets of scores for the held-out challenge
dataset, and average them to obtain one set of scores that was submitted.

%------------------------------------------------------------------------
\subsection{Domain Confusion}

% TODO Consult with Dr. Jacobs on whether this should be mentioned or not

This training strategy addresses the issue of dataset bias. It is apparent
that the three given datasets have disjoint domains. In addition, within
each dataset, data is skewed either towards mostly positive samples or
mostly negative ones. During training, it is very likely that the network
will cheat by identifying which dataset the clip is from, and make its
prediction based off of that information. In response, we take the audio
features near the end of the network, pass it through another convolutional
layer, and...

%------------------------------------------------------------------------
\subsection{Pseudo Labeling}

This technique is used by many contestants and attempted by us. The
assumption is that high-confidence predictions on the held-out challenge
dataset are accurate enough to be used as extra data for retraining or
fine-tuning. Basic procedures are:

\begin{alt_enumerate}
	\item{Train a model with only labeled data}
	\item{Predict scores of the held-out challenge dataset}
	\item{Find threshold where within it, predictions are accurate enough to
		be used as extra labels}
	\item{Add all predictions within threshold, and retrain or fine-tune
		network}
\end{alt_enumerate}

%------------------------------------------------------------------------
\section{Results}

Below are the final AUC scores for each of the 3 methods:\\

\begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|p{2cm}| }
	\hline
	\multicolumn{2}{|c|}{AUC Scores}\\
	\hline
	Method & AUC Score\\
	\hline
	Method 1 & 77.9\\
	Method 2 & ??.?\\
	Method 3 & 88.7\\
	\hline
\end{tabular}

%------------------------------------------------------------------------
\section{Future Work}

For method 1 and 2, no signal pre-processing techniques were deployed,
resulting in very noisy inputs to our network. It will be interesting to
find out if pre-processing will aid the prediction accuracy of our network.

%------------------------------------------------------------------------
\section{Conclusion}

Bird acoustic detection is an important task for many biological
applications, and the challenge has advanced this field in many ways. We
presented three methods, first two with a focus on unique network
architecture, and the third one as the embodiment of many years of machine
listening research by Thomas Grill. Many topics are still open for
research, like bird species identification and detection.

@@
